A Machine Learning Approach to MLB Catcher Framing

### 2021-12-10 Slides.Rmd
Add to slides

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(gbm)
library(randomForest)
knitr::opts_chunk$set(echo = FALSE)
```

@@

- Baseball catchers can influence the call of a ball or strike on how they catch the ball
- Some catchers are better than others at this skill
- Baseball teams are aware of this and are acquiring players good at this skill to win more games
- We want to quantify the best catcher's at framing for the 2021 season
- There are several factors that influence whether a pitch will be a strike or ball
- Catchers getting more strikes translates to more outs, and fewer runs for the opposing team

@@

```{r cars, echo = FALSE, message = FALSE, cache=TRUE}
load("Data/usable_data.Rda")
x=c(-0.95,0.95,0.95,-0.95,-0.95)
y=c(1.5,1.5,3.5,3.5,1.5)
sz=data.frame(x,y)
# Take some samples to explore
train = usable_data %>% slice(1000:150000) %>%
  mutate(strike = as.character(strike)) %>%
  select(-catcher_name)

@@

```{r}
x <- model.matrix(strike ~ ., data = data_no_catchers)[, -1]
y <- data_no_catchers$strike

@@

```{r, cache=TRUE}
#LASSO
lasso_fit <- glmnet(x[train, ], y[train], alpha = 1, lambda = lambda, family = 'binomial')
cv_lasso <- cv.glmnet(x[train, ], y[train], alpha = 1, family='binomial')

@@

non_zero_lasso_coef <- lasso_coeff[lasso_coeff != 0]
non_zero_lasso_coef
#Confusion matrix
lasso_mx <- table(pred=lasso_predict, true=y[test])
lasso_mx
error <- (lasso_mx[2] + lasso_mx[3])/length(lasso_predict)
1-error
```

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(gbm)
library(randomForest)
knitr::opts_chunk$set(echo = FALSE)
load("Data/usable_data.Rda")
```

@@

- Baseball catchers can influence the call of a ball or strike on how they catch the ball
- Some catchers are better than others at this skill
- Baseball teams are aware of this and are acquiring players good at this skill to win more games
- We want to quantify the best catcher's at framing for the 2021 season
- There are several factors that influence whether a pitch will be a strike or ball
- Catchers getting more strikes translates to more outs, and fewer runs for the opposing team
- https://baseballsavant.mlb.com/sporty-videos?playId=273b22a4-f522-4009-87d1-14176772182d

@@

```{r cars, echo = FALSE, message = FALSE, cache=TRUE}
x=c(-0.95,0.95,0.95,-0.95,-0.95)
y=c(1.5,1.5,3.5,3.5,1.5)
sz=data.frame(x,y)
# Take some samples to explore
train = usable_data %>% slice(1000:150000) %>%
  mutate(strike = as.character(strike)) %>%
  select(-catcher_name)

@@

```{r, cache=T}
x <- model.matrix(strike ~ ., data = data_no_catchers)[, -1]
y <- data_no_catchers$strike

@@

```{r, cache=TRUE, message=FALSE}
#LASSO
lasso_fit <- glmnet(x[train, ], y[train], alpha = 1, lambda = lambda, family = 'binomial')
cv_lasso <- cv.glmnet(x[train, ], y[train], alpha = 1, family='binomial')

@@

non_zero_lasso_coef <- lasso_coeff[lasso_coeff != 0]

#Confusion matrix
lasso_mx <- table(pred=lasso_predict, true=y[test])
lasso_mx
error <- (lasso_mx[2] + lasso_mx[3])/length(lasso_predict)
1-error
```

@@

## Model Comparison

- Boosting with interaction depth of 16, shrinkage of 0.1, and 500 trees has the highest accuracy

```{r}
df_rf %>%
  rename(parameter = nodesize) %>%
  bind_rows(final_df_gbm %>%
              group_by(trees) %>%
              summarize(accuracy = mean(accuracy)) %>%
              arrange(desc(accuracy)) %>%
              mutate(model = "boosting") %>%
              select(model, accuracy, trees) %>%
              rename(parameter = trees)) %>%
  arrange(desc(accuracy))
```

## Results

- To give credit to each catcher, we use this equation for every pitch:
  - If strike: 1 - Strike Probability
  - If ball: Strike Probability * -1
